Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-07-08 supplement.tex
Edits

\begin{smathpar}
\begin{array}{ll}
  \rulelabel{E-Foreach1} & \stg \vdash (\foreache{s}{\lambda y.\lambda
  z.c},\stl) \stepsto (\foreachr{\emptyset}{s}{\lambda y.\lambda z. c})\\
  \rulelabel{E-Foreach2} & \stg \vdash (\foreachr{s_1}{\{r\} \uplus s_2}{\lambda y.\lambda
  z.c},\stl) \stepsto ([r/z][s_1/y]c;\,\foreachr{s_1 \cup \{r\}}{s_2}{\lambda y.\lambda z. c})\\
  \rulelabel{E-Foreach3} & \stg \vdash (\foreachr{s}{\emptyset}{\lambda y.\lambda
  z.c},\stl) \stepsto (\cskip,\stl)\\
\end{array}
\end{smathpar}
%
\bigskip

@@

%
\textbf{Top-Level Reasoning} \quad \fbox {\(
\rg{I,R}{c}{G,I} \)}\\
%
%
\begin{minipage}{3.5in}
\rulelabel{RG-Txn}
\begin{smathpar}
\begin{array}{c}
\RULE
{
 \stable(R, I) \spc
  \stable(R,\I)\spc
  P(\stl,\stg) \Leftrightarrow \stl=\emptyset \wedge I(\stg)\\
  R_r(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_e(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc
  R_r \vdash \rg{P}{c}{Q}\\
  R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_c(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg')  \spc
  \stable(R_c,Q)\\
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow 
    G(\stg, \stl \gg \stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G \cup ID,I}
}
\end{array}
\end{smathpar}
\end{minipage}
%

%
\begin{minipage}{3.6in}
\rulelabel{RG-Par}
\begin{smathpar}
\begin{array}{c}
\RULE
{
 \rg{I,R \cup G_2 \cup ID}{t_1}{G_1 \cup ID,I} \\
 \rg{I,R \cup G_1 \cup ID}{t_2}{G_2 \cup ID,I}
}
{\rg{I,R}{t1||t2}{G_1\cup G_2 \cup ID,I}}
\end{array}
\end{smathpar}
\end{minipage}
%
%
\begin{minipage}{2in}
\rulelabel{RG-Conseq2}
\begin{smathpar}
\begin{array}{c}
\RULE
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G,I}\\
  \I' \Rightarrow \I \spc 
  R' \subseteq R \spc 
  \stable(R',\I')\\
  G \subseteq G' \spc
  \forall \stg,\stg'.~I(\stg) \wedge G'(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R'}{\ctxn{i}{\I'}{c}}{G',I}
}
\end{array}
\end{smathpar}
\end{minipage}
%

@@

In the sequence $\pi_1$, $R^{n_1}(\stg, \stg_1)$. By $I(\stg)$, $HI$ and Lemma 6.3, $I(\stg_1)$. By the rule E-Txn-Start, $t.\stl = \phi, t.\stg = \stg_1$ and $t.c = c$. Hence $P(t.\stl, \stg_1)$.

Expanding the definition of the assertion $Hc$ and instantiating it with $\stg = \stg_1$ and $\stg^{'} = \stg_2$, we would get $Q(t^{'}.\stl, \stg_2)$. However, the environment steps in run $\pi$ are in $R$, while the environment steps in assertion $Hc$ are in $\R_e$. Hence, we will now show that only environment steps in $\R_e$ can actually happen in the run $\pi$. We will show this in two steps. In the first step, we will prove that for all configurations $(t_p, \stg_p)$ in the run $\pi$ except possibly the last configuration, $\I_e(t_p.\stl, t_p.\stg, \stg_p)$.

@@

$(t^{'}, \stg_2) \rightarrow_{R}^{n_3 - 1} (t^{'}, \stg_{2}^{'}) \rightarrow (\texttt{SKIP}, \stg_3)$, where the first $n_3-1$ steps are only performed by the environment. Since the transaction commits at state $\stg_{2}^{'}$, by the E-Commit rule, $\I_c(t^{'}.\stl, \stg_2, \stg_{2}^{'})$. We will now show that all environment steps in the above run must be in $\R_c$. Again, we will show this in two steps. Let $m = n_3 - 1$ and $(t^{'}, \stg_2) \rightarrow_R (t^{'}, \stg_{21}) \rightarrow_R (t^{'}, \stg_{22}) \ldots \rightarrow_R (t^{'}, \stg_{2m}) \rightarrow (\texttt{SKIP}, \stg_3)$. We will show that $\I_c(t^{'}.\stl, \stg_2, \stg_{2k})$ for all $k, 1 \leq k \leq m$.

@@

By $HQ$, $Q(t^{'}.\stl, \stg_2)$ and Lemma 6.3 we have $Q(t^{'}.\stl, \stg_{2}^{'})$. Since all state changes so far have been made by the environment, $I(\stg_{2}^{'})$. By $HQG$, $G(\stg_{2}^{'}, t^{'}.\stl \gg \stg_{2}^{'})$. By the E-Commit rule, $\stg_3 = (t^{'}.\stl \gg \stg_{2}^{'})$. Hence, $G(\stg_{2}^{'}, \stg_3)$. All the steps of the transaction except the commit step do not change the global database state and by $HID$ belong to $G$. The commit step satisfies $G$. This proves the \texttt{step-guaranteed} assertion. Finally, by $HG$, $I(\stg_3)$.

@@

All the steps in $(\texttt{SKIP}, \stg_3) \rightarrow_{R}^{n_4} (\texttt{SKIP}, \stg^{'})$ are performed by the environment. Since $I(\stg_3)$, by $HI$ and Lemma 6.3, $I(\stg^{'})$.
\end{proof}

@@

\begin{itemize}
\item $(t,\stg) \rightarrow_{R}^{m_1} (t_{m}^{'}, \stg_1) \rightarrow (t_{m}, \stg_1)$, where $t_m.c = c2$. We denote this sequence as $\pi_1$.
\item $(t_m, \stg_1) \rightarrow_{R}^{m_2} (t_m, \stg_{1}^{'}) $ where all steps are taken by the environment. This sequence is denoted as $\pi_2$.
\item $(t_m, \stg_{1}^{'}) \rightarrow_{R}^{m_3} (t_{2}, \stg^{'}) \rightarrow (t^{'}, \stg^{'})$. This run is denoted as $\pi_3$.
\end{itemize}

@@

%There exists $s$ such that $s.c = c1$, $s.\sigma_l=t.\sigma_l$, $s.\sigma_v = t.\sigma_v$ and the run $(s, \sigma) \rightarrow (s_1, \sigma) \rightarrow_{R}^{m_1} (s_2, \sigma_1) \rightarrow (s^{'}, \sigma_1)$ with $s^{'}.c = \texttt{SKIP}$. Similarly, there exists $u$ such that $u.c = c2$, $u.\sigma_l = s^{'}.\sigma_l$, $u.\sigma_v = s^{'}.\sigma_v$ and the run $(u, \sigma_{1}^{'}) \rightarrow (u_1, \sigma_{1}^{'}) \rightarrow_{R}^{m_2} (u_2, \sigma^{'}) \rightarrow (u^{'}, \sigma^{'})$ with $R^{m_3}(\sigma_1, \sigma_{1}^{'})$ and $u^{'}.c = \texttt{SKIP}$. Also, $u^{'}.\sigma_l = t^{'}.\sigma_l$ and $u^{'}.\sigma_v = t^{'}.\sigma_v$.

@@

%\begin{lemma}
%If $\{\mathbb{I}, I, R\} t \{G, I\}$, then $\{\mathbb{I}, I, R \cup ID\} t \{G, I\}$, where $ID$ is the identity function
%\end{lemma}
%\begin{proof}
%Given $\sigma$ such that $I(\sigma)$ and the run $\pi_1$, $(t,\sigma) \rightarrow_{R \cup ID}^{n_1} (\texttt{SKIP}, \sigma^{'})$, we have to show (1) $I(\sigma^{'})$ and (2)$\texttt{commit-guarantee}(\mathbb{I}, R \cup ID, G, t, \sigma)$.
%
%We will transform the run $\pi_1$ to run $\pi_2$ of the form $(t,\sigma) \rightarrow_{R}^{n_2} (\texttt{SKIP}, \sigma^{'})$. Consider any sequence of two steps in the run $\pi_1$ of the form $(t_1, \sigma_1) \rightarrow_{R \cup ID} (t_1, \sigma_1) \rightarrow_{R \cup ID} (t_2, \sigma_2)$, i.e. the first step is taken by the environment in the form of the $ID$ function. Then we remove the first step and obtain the valid step $(t_1, \sigma_1) \rightarrow_{R \cup ID} (t_2, \sigma_2)$. We perform this transformation from left to right (i.e. increasing values of the step index) whenever the environment takes a step in the form of the $ID$ function. It is clear that the resulting run will only have environment steps in the form of $R$. Since the run $\pi_2$ also ends in $\sigma^{'}$, by the hypothesis $I(\sigma^{'})$. Also, the commit step in the run $\pi_2$ will involve the same states as the commit step in $\pi_1$. Hence, $\texttt{commit-guarantee}(\mathbb{I}, R \cup ID, G, t, \sigma)$. 
%\end{proof}

@@

Consider $\stg$ such that $I(\stg)$, and let $(t_1 || t_2, \stg) \rightarrow_{R}^{n} (\texttt{SKIP}, \stg^{'})$. We have to show (1) $I(\stg^{'})$ and (2) $\texttt{step-guaranteed}( R, G_1 \cup G_2 \cup ID, t1 || t_2, \stg)$.

Suppose that $t_1$ commits before $t_2$ in the execution sequence. Consider the sequence upto (and including) the commit step of $t_1$, i.e. $(t_1 || t_2, \stg) \rightarrow_{R}^{n_1} (t_{1}^{'} || t_{2}^{'}, \stg_1) \rightarrow (t_{2}^{'}, \stg_{1}^{'})$. In this sequence, all steps apart from the steps taken by $t_1$ belong to $R \cup ID$, since any step taken by $t_2$ cannot change the global database state. Hence, there exists the sequence $(t_1, \stg) \rightarrow_{R \cup ID}^{n_1} (t_{1}^{'}, \stg_1) \rightarrow (\texttt{SKIP}, \stg_{1}^{'})$. Since $R \cup ID \subseteq R \cup G_2 \cup ID$, by H1 and Lemma, $I(\stg_{1}^{'})$ and $G_1(\stg_1, \stg_{1}^{'})$. Now, consider the entire run from the perspective of $t_2$. All steps taken by $t_1$ except the commit step do not change the global database state, and the change during the commit step belongs to $G_1$. Hence, all steps in the run apart from the steps taken by $t_2$ belong to $R \cup G_1 \cup ID$. Hence, there exists a run $(t_2, \stg) \rightarrow_{R \cup G_1 \cup ID}^{n} (\texttt{SKIP}, \stg^{'})$. By H2 $I(\stg^{'})$.

>>>>

%\begin{minipage}
\begin{smathpar}
\begin{array}{ll}
  \rulelabel{E-Foreach1} & \stg \vdash (\foreache{s}{\lambda y.\lambda
  z.c},\stl) \stepsto (\foreachr{\emptyset}{s}{\lambda y.\lambda z. c})\\
  \rulelabel{E-Foreach2} & \stg \vdash (\foreachr{s_1}{\{r\} \uplus s_2}{\lambda y.\lambda
  z.c},\stl) \stepsto ([r/z][s_1/y]c;\,\foreachr{s_1 \cup \{r\}}{s_2}{\lambda y.\lambda z. c})\\
  \rulelabel{E-Foreach3} & \stg \vdash (\foreachr{s}{\emptyset}{\lambda y.\lambda
  z.c},\stl) \stepsto (\cskip,\stl)\\
\end{array}
\end{smathpar}
%
%\bigskip

@@

%
\textbf{Top-Level Reasoning} \quad \fbox {\(
\rg{I,R}{c}{G,I} \)}\\
%
%
\begin{minipage}{3.5in}
\rulelabel{RG-Txn}
\begin{smathpar}
\begin{array}{c}
\RULE
{
 \stable(R, I) \spc
  \stable(R,\I)\spc
  P(\stl,\stg) \Leftrightarrow \stl=\emptyset \wedge I(\stg)\\
  \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_e(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc
  \R_e \vdash \rg{P}{c}{Q}\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1. \I_c(\stl, \stg_1, \stg) \wedge R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg')  \spc
  \stable(\R_c,Q)\\
  \forall \stl,\stg.~ Q(\stl,\stg) \Rightarrow 
    G(\stg, \stl \gg \stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G \cup ID,I}
}
\end{array}
\end{smathpar}
\end{minipage}
%

%
\begin{minipage}{3.6in}
\rulelabel{RG-Par}
\begin{smathpar}
\begin{array}{c}
\RULE
{
 \rg{I,R \cup G_2 \cup ID}{t_1}{G_1 \cup ID,I} \\
 \rg{I,R \cup G_1 \cup ID}{t_2}{G_2 \cup ID,I}
}
{\rg{I,R}{t_1||t_2}{G_1\cup G_2 \cup ID,I}}
\end{array}
\end{smathpar}
\end{minipage}
%
%
\begin{minipage}{2in}
\rulelabel{RG-Conseq2}
\begin{smathpar}
\begin{array}{c}
\RULE
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G,I}\\
  \I' \Rightarrow \I \spc 
  R' \subseteq R \spc 
  \stable(R',\I')\\
  G \subseteq G' \spc
  \forall \stg,\stg'.~I(\stg) \wedge G'(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R'}{\ctxn{i}{\I'}{c}}{G',I}
}
\end{array}
\end{smathpar}
\end{minipage}
%

@@

In the sequence $\pi_1$, $R^{n_1}(\stg, \stg_1)$. By $I(\stg)$, $HI$ and Lemma 3.3, $I(\stg_1)$. By the rule E-Txn-Start, $t.\stl = \phi, t.\stg = \stg_1$ and $t.c = c$. Hence $P(t.\stl, \stg_1)$.

Expanding the definition of the assertion $Hc$ and instantiating it with $\stg = \stg_1$ and $\stg^{'} = \stg_2$, we would get $Q(t^{'}.\stl, \stg_2)$. However, the environment steps in sequence $\pi$ are in $R$, while the environment steps in assertion $Hc$ are in $\R_e$. Hence, we will now show that only environment steps in $\R_e$ can actually happen in the sequence $\pi$. We will show this in two steps. In the first step, we will prove that for all configurations $(t_p, \stg_p)$ in the sequence $\pi$ except possibly the last configuration, $\I_e(t_p.\stl, t_p.\stg, \stg_p)$.

@@

$(t^{'}, \stg_2) \rightarrow_{R}^{n_3 - 1} (t^{'}, \stg_{2}^{'}) \rightarrow (\texttt{SKIP}, \stg_3)$, where the first $n_3-1$ steps are only performed by the environment. Since the transaction commits at state $\stg_{2}^{'}$, by the E-Commit rule, $\I_c(t^{'}.\stl, \stg_2, \stg_{2}^{'})$. We will now show that all environment steps in the above sequence must be in $\R_c$. Again, we will show this in two steps. Let $m = n_3 - 1$ and $(t^{'}, \stg_2) \rightarrow_R (t^{'}, \stg_{21}) \rightarrow_R (t^{'}, \stg_{22}) \ldots \rightarrow_R (t^{'}, \stg_{2m}) \rightarrow (\texttt{SKIP}, \stg_3)$. We will show that $\I_c(t^{'}.\stl, \stg_2, \stg_{2k})$ for all $k, 1 \leq k \leq m$.

@@

By $HQ$, $Q(t^{'}.\stl, \stg_2)$ and Lemma 3.3 we have $Q(t^{'}.\stl, \stg_{2}^{'})$. Since all state changes so far have been made by the environment, $I(\stg_{2}^{'})$. By $HQG$, $G(\stg_{2}^{'}, t^{'}.\stl \gg \stg_{2}^{'})$. By the E-Commit rule, $\stg_3 = (t^{'}.\stl \gg \stg_{2}^{'})$. Hence, $G(\stg_{2}^{'}, \stg_3)$. All the steps of the transaction except the commit step do not change the global database state and by $HID$ belong to $G$. The commit step satisfies $G$. This proves the \texttt{step-guaranteed} assertion. Finally, by $HG$, $I(\stg_3)$.

@@

All the steps in $(\texttt{SKIP}, \stg_3) \rightarrow_{R}^{n_4} (\texttt{SKIP}, \stg^{'})$ are performed by the environment. Since $I(\stg_3)$, by $HI$ and Lemma 3.3, $I(\stg^{'})$.
\end{proof}

@@

\begin{itemize}
\item $(t,\stg) \rightarrow_{R}^{m_1} (t_{m}^{'}, \stg_1) \rightarrow (t_{m}, \stg_1)$, where $t_m.c = c2$. We denote this sequence as $\pi_1$.
\item $(t_m, \stg_1) \rightarrow_{R}^{m_2} (t_m, \stg_{1}^{'}) $ where all steps are taken by the environment. This sequence is denoted as $\pi_2$.
\item $(t_m, \stg_{1}^{'}) \rightarrow_{R}^{m_3} (t_{2}, \stg^{'}) \rightarrow (t^{'}, \stg^{'})$. This sequence is denoted as $\pi_3$.
\end{itemize}

@@

%There exists $s$ such that $s.c = c1$, $s.\sigma_l=t.\sigma_l$, $s.\sigma_v = t.\sigma_v$ and the sequence $(s, \sigma) \rightarrow (s_1, \sigma) \rightarrow_{R}^{m_1} (s_2, \sigma_1) \rightarrow (s^{'}, \sigma_1)$ with $s^{'}.c = \texttt{SKIP}$. Similarly, there exists $u$ such that $u.c = c2$, $u.\sigma_l = s^{'}.\sigma_l$, $u.\sigma_v = s^{'}.\sigma_v$ and the sequence $(u, \sigma_{1}^{'}) \rightarrow (u_1, \sigma_{1}^{'}) \rightarrow_{R}^{m_2} (u_2, \sigma^{'}) \rightarrow (u^{'}, \sigma^{'})$ with $R^{m_3}(\sigma_1, \sigma_{1}^{'})$ and $u^{'}.c = \texttt{SKIP}$. Also, $u^{'}.\sigma_l = t^{'}.\sigma_l$ and $u^{'}.\sigma_v = t^{'}.\sigma_v$.

@@

%\begin{lemma}
%If $\{\mathbb{I}, I, R\} t \{G, I\}$, then $\{\mathbb{I}, I, R \cup ID\} t \{G, I\}$, where $ID$ is the identity function
%\end{lemma}
%\begin{proof}
%Given $\sigma$ such that $I(\sigma)$ and the sequence $\pi_1$, $(t,\sigma) \rightarrow_{R \cup ID}^{n_1} (\texttt{SKIP}, \sigma^{'})$, we have to show (1) $I(\sigma^{'})$ and (2)$\texttt{commit-guarantee}(\mathbb{I}, R \cup ID, G, t, \sigma)$.
%
%We will transform the sequence $\pi_1$ to sequence $\pi_2$ of the form $(t,\sigma) \rightarrow_{R}^{n_2} (\texttt{SKIP}, \sigma^{'})$. Consider any sequence of two steps in the sequence $\pi_1$ of the form $(t_1, \sigma_1) \rightarrow_{R \cup ID} (t_1, \sigma_1) \rightarrow_{R \cup ID} (t_2, \sigma_2)$, i.e. the first step is taken by the environment in the form of the $ID$ function. Then we remove the first step and obtain the valid step $(t_1, \sigma_1) \rightarrow_{R \cup ID} (t_2, \sigma_2)$. We perform this transformation from left to right (i.e. increasing values of the step index) whenever the environment takes a step in the form of the $ID$ function. It is clear that the resulting sequence will only have environment steps in the form of $R$. Since the sequence $\pi_2$ also ends in $\sigma^{'}$, by the hypothesis $I(\sigma^{'})$. Also, the commit step in the sequence $\pi_2$ will involve the same states as the commit step in $\pi_1$. Hence, $\texttt{commit-guarantee}(\mathbb{I}, R \cup ID, G, t, \sigma)$. 
%\end{proof}

@@

Consider $\stg$ such that $I(\stg)$, and let $(t_1 || t_2, \stg) \rightarrow_{R}^{n} (\texttt{SKIP}, \stg^{'})$. We have to show (1) $I(\stg^{'})$ and (2) $\texttt{step-guaranteed}( R, G_1 \cup G_2 \cup ID, t_1 || t_2, \stg)$.

Suppose that $t_1$ commits before $t_2$ in the execution sequence. Consider the sequence upto (and including) the commit step of $t_1$, i.e. $(t_1 || t_2, \stg) \rightarrow_{R}^{n_1} (t_{1}^{'} || t_{2}^{'}, \stg_1) \rightarrow (t_{2}^{'}, \stg_{1}^{'})$. In this sequence, all steps apart from the steps taken by $t_1$ belong to $R \cup ID$, since any step taken by $t_2$ cannot change the global database state. Hence, there exists the sequence $(t_1, \stg) \rightarrow_{R \cup ID}^{n_1} (t_{1}^{'}, \stg_1) \rightarrow (\texttt{SKIP}, \stg_{1}^{'})$. Since $R \cup ID \subseteq R \cup G_2 \cup ID$, by H1 and Lemma 3.13, $I(\stg_{1}^{'})$ and $G_1(\stg_1, \stg_{1}^{'})$. Now, consider the entire sequence from the perspective of $t_2$. All steps taken by $t_1$ except the commit step do not change the global database state, and the change during the commit step belongs to $G_1$. Hence, all steps in the sequence apart from the steps taken by $t_2$ belong to $R \cup G_1 \cup ID$. Hence, there exists a sequence $(t_2, \stg) \rightarrow_{R \cup G_1 \cup ID}^{n} (\texttt{SKIP}, \stg^{'})$. By H2 $I(\stg^{'})$.
